---
title: "Machine Learning Exercise - Gym Movements"
author: "Charlie Becker"
date: "4/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways (labeled A-E). More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Load libraries and data

The 'foreach' and 'doMC' packages were loaded to utilize 3of 4 computing cores to speed up processing for this dataset.  

The data contained nearly 20,000 observations of 160 total variables from a csv file, though many variables had significant amounts of NA values, missing or blank values, and error values by trying to divide by zero.  To begin cleaning the data - for both the test and train - all missing and error values were converted to NA's and then all NA values summed across columns/variables.  Any column that had a majority of NA's was discarded.  Lastly, all columns that were not relevant predictors (such as names/times) were removed, leaving us with a total of 52 predictor variables.  

Since the dataset is relatively large, and computational time will be on the order of hours - a straight randomized 50/50 split of the "training" set will be used for cross validation.

```{r eval = FALSE}
library(caret)
library(foreach)
library(doMC)
registerDoMC(cores = 3)

trn <- read.csv("/Users/charlesbecker/Downloads/pml-training.csv",na.strings=c("NA","#DIV/0!",""))
tst <- read.csv("/Users/charlesbecker/Downloads/pml-testing.csv", na.strings=c("NA","#DIV/0!",""))

sumNA <- function(x) sum(is.na(x))

NA_sum_trn <- apply(trn, 2, sumNA)
NA_sum_tst <- apply(tst, 2, sumNA)

NA_row_train <- which(NA_sum_trn > nrow(trn)*.5)
NA_row_test <- which(NA_sum_tst > nrow(tst)*.5)

training <- trn[,-NA_row_train]
testing <- tst[,-NA_row_test]

training <- training[,-(1:7)]
testing <- testing[,-(1:7)]

samp <- sample(nrow(training), nrow(training)/2)
new_train <- training[samp,]
cv_data <- training[-samp,]
```

### Build the models

A varity of models will be trained on the same training data and tested on the cross validation set to see error rates.  Classification models included are:  Rpart decision tree, Naive Bayes, Stochastic Gradient Boosting, Neural network, Logistical Regression Boosting, K-Nearest Neighbor, and Random Forest.  All models will be run under the "caret" package wrapper.

```{r eval = FALSE}
mod_rpart <- train(classe ~ ., method = "rpart", data = new_train)
mod_nb <- train(classe ~ ., method = "nb", data = new_train)
mod_gbm <- train(classe ~ ., method = "gbm", data = new_train, verbose = F)
mod_nnet <- train(classe ~ ., method = "nnet", data = new_train)
mod_logit <- train(classe ~ ., method = "LogitBoost", data = new_train)
mod_knn <- train(classe ~ ., method = "knn", data = new_train)
mod_rf <- train(classe ~ ., method = "rf", data = new_train, prox = T)
```

### Predict on the cross validation data and report error rates

```{r}
table(p_rpart, cv_data$classe)
confusionMatrix(p_rpart, cv_data$classe)$overall[1]
table(p_nb, cv_data$classe)
confusionMatrix(p_nb, cv_data$classe)$overall[1]
table(p_nnet, cv_data$classe)
confusionMatrix(p_nnet, cv_data$classe)$overall[1]
table(p_gbm, cv_data$classe)
confusionMatrix(p_gbm, cv_data$classe)$overall[1]
table(p_logit, cv_data$classe)
confusionMatrix(p_logit, cv_data$classe)$overall[1]
table(p_knn, cv_data$classe)
confusionMatrix(p_knn, cv_data$classe)$overall[1]
table(p_rf, cv_data$classe)
confusionMatrix(p_rf, cv_data$classe)$overal[1]
```

### Analyze results

There's a wide variety of accuracy between the models ranging from ~ 40-99 %. The random forest was the most accurate with the ~ 1% error rate.  Since the random forest did a better job with all 5 classifiers than any other model, an ensemble stacking approach was not used. However, the Gradient Boosting model can provide some insight on the influence of the variables as show below.

```{r mod_gbm, echo = FALSE}
summary(mod_gbm)
```

### Use model to predict on 20 observation test set

```{r}
predict(mod_rf, tst)
```


